---
title: "Q(2)"
author: "luhuan wu"
date: "4/12/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Data Analysis
```{r}
ts2 <- ts(unlist(read.csv("/Users/leah/Downloads/q2_train.csv", as.is = TRUE)))
```

```{r}
plot(ts2, type = 'l')
ts2_true<-as.numeric( ts2[526:1050] ) # select effective data
plot(ts2_true,type = "l")
ts2_positive<-ts2_true+2  # shift upwards to get the positive value so we can take log transformation
plot(ts2_positive,type = "l")
```
##Transform the data to stabilize the variance.
```{r}
ts2.log <- log(ts2_positive)
plot(ts2.log, type ='l')
```
##Try taking square root to stablize the variance
```{r}
ts2.sqrt <- sqrt(ts2_positive)
plot(ts2.sqrt, type ='l')
```
We can observe an upward trend. Use differencing to detrend.

```{r}
ts2.sqrt.d <- diff(ts2.sqrt)
plot(ts2.sqrt.d, type = 'l')
```
##Remove the mean structure
#differencing
```{r}
ts2.sqrt.dd <- diff(ts2.sqrt.d)
plot(ts2.sqrt.dd, type = 'l')
```

##Remove the mean structure
#differencing
```{r}
ts2.log.d <- diff(ts2.log)
plot(ts2.log.d, type = 'l')
```
#try differencing again
```{r}
ts2.log.dd <- diff(ts2.log.d)
plot(ts2.log.dd, type = 'l')
```

## Check out the performance for differenced data and compare.
```{r}
acf(ts2.log.d, lag.max = 100)
acf(ts2.log.dd, lag.max = 100)
pacf(ts2.log.d, lag.max = 100)
pacf(ts2.log.dd,lag.max = 100)
```
No big improvments from 1st-order differencing. So we use 1st-order differenced data. 
TODO: try 2nd-order differencing later.

Enlarge the lag.max to observe whether spkikes occur with periodicity.
```{r}
acf(ts2.log.d, lag.max = 500)
acf(ts2.log.dd, lag.max = 500)
pacf(ts2.log.d, lag.max = 500)
pacf(ts2.log.dd,lag.max = 500)
```

Seasonality occurs at lags multiple of 52.

##Remove the seasonality
```{r}
ts2.log.dD <- diff(ts2.log.d, 52)
plot(ts2.log.dd, type = 'l')
acf(ts2.log.dD,lag.max = 500)
pacf(ts2.log.dD, lag.max = 500)
```

(p,d,q) x (P,D,Q)_s
From the acf and pacf above, we set d = 1, D = 1, p = P = 0, s = 52
try q = 2, Q = 2, 3, 4

## Model Fitting 

We fit the models with ARIMA(0,1,q)x(0,1,Q)_52 with q,Q taking values over(2,3,4)x(5,6,7).

```{r}
m2_22 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(0, 1, 2), period = 52))
m2_22
tsdiag(m2_22)
AIC(m2_22)
BIC(m2_22)
#AIC=-830.4667
#BIC = -809.6818
#p values at lag larger than 1 is too below 0.05.
```


```{r}
m2_23 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(0, 1, 3), period = 52))
m2_23
tsdiag(m2_23)
AIC(m2_23)
BIC(m2_23)
#AIC = -828.4839
#BIC =  -803.542
```

```{r}
m2_24 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(0, 1, 4), period = 52))
m2_24
tsdiag(m2_24)
AIC(m2_24)
BIC(m2_24)
#AIC=-482.163
#-457.2211
```

The above models are not reasonable considering their p values. Now we try adding AR components.

```{r}
m2_0213 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(1, 1, 3), period = 52))
m2_0213
```
```{r}
tsdiag(m2_0213)
```
```{r}
m2_0214 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(1, 1, 4), period = 52))
m2_0214
```
```{r}
tsdiag(m2_0214)
AIC(m2_0214)
BIC(m2_0214)
```
```{r}
m2_0223 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(2, 1, 3), period = 52))
m2_0223
tsdiag(m2_0223)
AIC(m2_0223)
BIC(m2_0223)
#AIC = -824.6507
#BIC = -791.3949
```

```{r}
m2_0224 <- arima(ts2.log, order = c(0, 1, 2), seasonal = list(order = c(2, 1, 4), period = 52))
m2_0224
tsdiag(m2_0224)
AIC(m2_0224)
BIC(m2_0224)
```

```{r}
m2_1213 <- arima(ts2.log, order = c(1, 1, 2), seasonal = list(order = c(1, 1, 3), period = 52))
m2_1213
tsdiag(m2_1213)
AIC(m2_1213)
BIC(m2_1213)
```

```{r}
m2_1214 <- arima(ts2.log, order = c(1, 1, 2), seasonal = list(order = c(1, 1, 4), period = 52))
m2_1214
tsdiag(m2_1214)
AIC(m2_1214)
BIC(m2_1214)
```
## Cross Validation
We use the cross validation function defined in Q1(1)-copy.Rmd

```{r}
MSE2_26 <- computeCVmse(c(0,1, 2), c(0,1,6))
MSE36 <- computeCVmse(c(0, 1, 3), c(0,1,6))
MSE46 <- computeCVmse(c(0, 1, 4), c(0,1,6))


MSE26# 0.2112802 0.1293311 0.5460744 0.1423115 0.1436121
MSE36# 0.2043168 0.1445004 0.5460334 0.1428842 0.1530916
MSE46#0.2039268 0.1424566 0.5468994 0.1417810 0.1530116
```