---
title: "q5"
author: "luhuan wu"
date: "4/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Compute_MSE function for cross-validation
```{r}
computeCVmse <- function(order.totry, seasorder.totry){
  MSE=numeric()
  len=length(ts.log)
  for(k in 5:1){
    train.dt=ts.log[1:(len-52*k)]
    test.dt=ts.log[(len-52*k+1):(len-52*(k-1))]
    mod=arima(train.dt,order=order.totry,
              seasonal=list(order=seasorder.totry,period=52))
    fcst=predict(mod,n.ahead=52)
    MSE[k]=mean((exp(fcst$pred) - exp(test.dt))^2)
    
  }
  return(MSE)
}
```
## Exploratory Data Analysis
```{r}
ts5 <- ts(unlist(read.csv("/Users/leah/Downloads/q5_train.csv", as.is = TRUE)))
ts5_data<-as.numeric( ts5[526:1050] )
plot(ts5_data,type = "l")
ts5<-ts(ts5_data,frequency = 52)
plot(ts5_data, main = "ts5_data", type = "l")
ts5_pos <- ts5 + 2
ts5.log <- log(ts5_pos)
plot(ts5.log, main = "ts5.log")
ts5.sqrt <- sqrt(ts5_pos)
plot(ts5.sqrt, main = "ts5.sqrt")

acf(ts5.log, lag.max = 100, main = "ts5.log")
acf(ts5.log, lag.max = 500, main = "ts5.log")

pacf(ts5.log, lag.max = 100, main = "ts5.log")
pacf(ts5.log, lag.max = 500, main = "ts5.log")


```
##Use auto.arima
```{r}
library(forecast)
m_try1 = auto.arima(ts5)
```
#diag
```{r}
m_try1
tsdiag(m_try1)
```

##Use robustarima
```{r}
m_try2 <- arima.rob(ts5 ~ 1, p=2, d=1)
m_try2
```
##Try differencing
```{r}
ts5d <- diff(ts5)
ts5dD <- diff(ts5d, 52)
plot(ts5d)
acf(ts5d, lag.max = 104)
pacf(ts5d, lag.max = 500)
acf(ts5d, lag.max = 500)
plot(ts5dD)
acf(ts5dD, lag.max = 104)
pacf(ts5dD, lag.max = 500)
acf(ts5dD, lag.max = 500)

```

##Fit models and forcasting

auto.arima gives ARIMA(0,1,2)(0,0,2)[52]  

```{r}
m5.012003 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(0, 0, 3), period = 52))
m5.012012 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(0, 1, 2), period = 52))
m5.012102 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(1, 0, 2), period = 52))
m5.012103 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(1, 0, 3), period = 52))
m5.112002 <- arima(ts5, order = c(1, 1, 2), seasonal = list(order = c(0, 0, 2), period = 52))
m5.113002 <- arima(ts5, order = c(1, 1, 3), seasonal = list(order = c(0, 0, 2), period = 52))
```

```{r}
m5.112002 <- arima(ts5, order = c(1, 1, 2), seasonal = list(order = c(0, 0, 2), period = 52))
```

```{r}
m5.113002 <- arima(ts5, order = c(1, 1, 3), seasonal = list(order = c(0, 0, 2), period = 52))
```

```{r}
pred.012003 <- predict(m5.012003, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.012003)), c(ts5, pred.012003), type = 'l', col = 1, main = "m5.012003")
points((length(ts5) + 1) : (length(ts5) + length(pred.012003)), pred.012003, type = 'l', col = 2)
pred.012003 <- predict(m5.012003, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.012003)), c(ts5, pred.012003), type = 'l', col = 1, main = "m5.012003")
points((length(ts5) + 1) : (length(ts5) + length(pred.012003)), pred.012003, type = 'l', col = 2)

pred.012012 <- predict(m5.012012, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.012012)), c(ts5, pred.012012), type = 'l', col = 1, main = "m5.012012")
points((length(ts5) + 1) : (length(ts5) + length(pred.012012)), pred.012012, type = 'l', col = 2)

pred.012102 <- predict(m5.012102, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.012102)), c(ts5, pred.012102), type = 'l', col = 1, main = "m5.012102")
points((length(ts5) + 1) : (length(ts5) + length(pred.012102)), pred.012102, type = 'l', col = 2)

pred.012103 <- predict(m5.012103, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.012103)), c(ts5, pred.012103), type = 'l', col = 1, main = "m5.012103")
points((length(ts5) + 1) : (length(ts5) + length(pred.012103)), pred.012103, type = 'l', col = 2)

pred.112002 <- predict(m5.112002, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.112002)), c(ts5, pred.112002), type = 'l', col = 1, main = "m5.112002")
points((length(ts5) + 1) : (length(ts5) + length(pred.112002)), pred.112002, type = 'l', col = 2)

pred.113002 <- predict(m5.113002, n.ahead = 208 )$pred
plot(1:(length(ts5) + length(pred.113002)), c(ts5, pred.113002), type = 'l', col = 1, main = "m5.112003")
points((length(ts5) + 1) : (length(ts5) + length(pred.113002)), pred.113002, type = 'l', col = 2)

```

##Compute MSE
```{r}
computeCVmse <- function(order.totry, seasorder.totry){
  MSE=numeric()
  len=length(ts5_data)
  for(k in 5:1){
    train.dt=ts5_data[1:(len-52*k)]
    test.dt=ts5_data[(len-52*k+1):(len-52*(k-1))]
    mod=arima(train.dt,order=order.totry,
              seasonal=list(order=seasorder.totry,period=52))
    fcst=predict(mod,n.ahead=52)
    MSE[k]=mean((exp(fcst$pred) - exp(test.dt))^2)
    
  }
  return(MSE)
}
```

##Diagonistic
```{r}
m5.012002 <- m_try1

m5.012002
m5.012003
m5.012012
m5.012102
m5.012103
m5.112002
m5.113002

tsdiag(m5.012002)
tsdiag(m5.012003)
tsdiag(m5.012012)
tsdiag(m5.012102)
tsdiag(m5.012103)
tsdiag(m5.112002)
tsdiag(m5.113002)

AIC(m5.012002)
AIC(m5.012003)
AIC(m5.012012)
AIC(m5.012102)
AIC(m5.012103)
AIC(m5.112002)
AIC(m5.113002)

BIC(m5.012002)
BIC(m5.012003)
BIC(m5.012012)
BIC(m5.012102)
BIC(m5.012103)
BIC(m5.112002)
BIC(m5.113002)

```
```{r}
MSE5_012002 
MSE5_012003
MSE5_012012
MSE5_012102
MSE5_012103
MSE5_112002 
MSE5_113002 
```

```{r}
MSE5_012003 <- computeCVmse(c(0, 1, 2), c(0,0,3))
MSE5_012012 <- computeCVmse(c(0, 1, 2), c(0,1,2))
MSE5_012102 <- computeCVmse(c(0, 1, 2), c(1,0,2))
MSE5_012103<- computeCVmse(c(0, 1, 2), c(1,0,3))
MSE5_112002 <- computeCVmse(c(1, 1, 2), c(0,0,2))
MSE5_113002 <- computeCVmse(c(1, 1, 3), c(0,0,2))
```
```{r}
MSE5_012002 <- computeCVmse(c(0, 1, 2), c(0, 0, 2))
```
## Forecasting

```{r}
pred_try1 <- predict(m_try1, n.ahead = 104 )$pred
plot(1:(length(ts5) + length(pred_try1)), c(ts5, pred_try1), type = 'l', col = 1)
points((length(ts5) + 1) : (length(ts5) + length(pred_try1)), pred_try1, type = 'l', col = 2)
```

### Fit modelling part II
MSE5_012102
MSE5_012103
MSE5_112002 
```{r}
m5.013102 <- arima(ts5, order = c(0, 1, 3), seasonal = list(order = c(1, 0, 2), period = 52))
m5.014102 <- arima(ts5, order = c(0, 1, 4), seasonal = list(order = c(1, 0, 2), period = 52))
m5.012202 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(2, 0, 2), period = 52))

m5.012203 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(2, 0, 3), period = 52))
m5.112103 <- arima(ts5, order = c(1, 1, 2), seasonal = list(order = c(1, 0, 3), period = 52))
m5.013103 <- arima(ts5, order = c(0, 1, 3), seasonal = list(order = c(1, 0, 3), period = 52))

m5.112102 <- arima(ts5, order = c(1, 1, 2), seasonal = list(order = c(1, 0, 2), period = 52))
m5.112101 <- arima(ts5, order = c(1, 1, 2), seasonal = list(order = c(1, 0, 1), period = 52))
```

```{r}
MSE5_013102 <- computeCVmse(c(0, 1, 3), c(1, 0, 2))
#MSE5_013102 <- computeCVmse(c(0, 1, 3), c(1, 0, 2))
MSE5_012202 <- computeCVmse(c(0, 1, 2), c(2,0,2))
MSE5_112103<- computeCVmse(c(1, 1, 2), c(1, 0, 3))
MSE5_013103 <- computeCVmse(c(0, 1, 3), c(1,0,3))
MSE5_112102 <- computeCVmse(c(1, 1, 2), c(1,0,2))
MSE5_112101 <- computeCVmse(c(1, 1, 2), c(1,0,1))
```
```{r}
MSE5_014102 <- computeCVmse(c(0, 1, 4), c(1, 0, 2))
```
```{r}
MSE5_012203 <- computeCVmse(c(0, 1, 2), c(2, 0, 3))
```

##Diagonistics

```{r}
m5.013102
m5.014102
m5.012202
m5.012203
m5.112103
m5.013103
m5.112002
m5.112101

tsdiag(m5.012002)
tsdiag(m5.012003)
tsdiag(m5.012012)
tsdiag(m5.012203)
tsdiag(m5.112103)
tsdiag(m5.013103)
tsdiag(m5.112102)
tsdiag(m5.112101)

print("AIC")
AIC(m5.012002)
AIC(m5.012003)
AIC(m5.012012)
AIC(m5.012203)
AIC(m5.112103)
AIC(m5.013103)
AIC(m5.112102)
AIC(m5.112101)

print("BIC")
BIC(m5.012002)
BIC(m5.012003)
BIC(m5.012012)
BIC(m5.012203)
BIC(m5.112103)
BIC(m5.013103)
BIC(m5.112102)
BIC(m5.112101)

print("MSE")
MSE5_013102 
MSE5_014102 
MSE5_012202 
MSE5_012203
MSE5_112103
MSE5_013103 
MSE5_112102
MSE5_112101 
```

##Forecasting

```{r}
pred5.013102 <- predict(m5.013102, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.013102)), c(ts5, pred5.013102), type = 'l', col = 1, main = "m5.013102")
points((length(ts5) + 1) : (length(ts5) + length(pred5.013102)), pred5.013102, type = 'l', col = 2)

pred5.014102 <- predict(m5.014102, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.014102)), c(ts5, pred5.014102), type = 'l', col = 1, main = "m5.014102")
points((length(ts5) + 1) : (length(ts5) + length(pred5.014102)), pred5.014102, type = 'l', col = 2)

pred5.012202 <- predict(m5.012202, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.012202)), c(ts5, pred5.012202), type = 'l', col = 1, main = "m5.012202")
points((length(ts5) + 1) : (length(ts5) + length(pred5.012202)), pred5.012202, type = 'l', col = 2)

pred5.012203 <- predict(m5.012203, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.012203)), c(ts5, pred5.012203), type = 'l', col = 1, main = "m5.012203")
points((length(ts5) + 1) : (length(ts5) + length(pred5.012203)), pred5.012203, type = 'l', col = 2)

pred5.112103 <- predict(m5.112103, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.112103)), c(ts5, pred5.112103), type = 'l', col = 1, main = "m5.112103")
points((length(ts5) + 1) : (length(ts5) + length(pred5.112103)), pred5.112103, type = 'l', col = 2)

pred5.013103 <- predict(m5.013103, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.013103)), c(ts5, pred5.013103), type = 'l', col = 1, main = "m5.013103")
points((length(ts5) + 1) : (length(ts5) + length(pred5.013103)), pred5.013103, type = 'l', col = 2)

pred5.112102 <- predict(m5.112102, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.112102)), c(ts5, pred5.112102), type = 'l', col = 1, main = "m5.112102")
points((length(ts5) + 1) : (length(ts5) + length(pred5.112102)), pred5.112102, type = 'l', col = 2)

pred5.112101 <- predict(m5.112101, n.ahead = 104)$pred
plot(1:(length(ts5) + length(pred5.112101)), c(ts5, pred5.112101), type = 'l', col = 1, main = "m5.112101")
points((length(ts5) + 1) : (length(ts5) + length(pred5.112101)), pred5.112102, type = 'l', col = 2)

```

##Try seasonal diffrencing
add seasonal differencing component to the models below by 1
012102
012103
013103
```{r}
m5.012112 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(1, 1, 2), period = 52))
m5.012113 <- arima(ts5, order = c(0, 1, 2), seasonal = list(order = c(1, 1, 3), period = 52))
##m5.013113 <- arima(ts5, order = c(0, 1, 3), seasonal = list(order = c(1, 1, 3), period = 52))
```

##Compute MSE
```{r}
m5.013113 <- arima(ts5, order = c(0, 1, 3), seasonal = list(order = c(1, 1, 3), period = 52))
MSE5_012112 <- computeCVmse(c(0, 1, 2), c(1, 1, 2))
MSE5_012113 <- computeCVmse(c(0, 1, 2), c(1, 1, 3))
MSE5_013113 <- computeCVmse(c(0, 1, 3), c(1, 1, 3))
```